Why Real-World Subgraph Isomorphism is Easy, \\ and What This Means for Graph Databases

### 2016-10-11 paper/paper.tex
Number some equations

We can approximately predict the location of the phase transition by calculating (with
simplifications regarding rounding and independence) the expected number of solutions for given
parameters. Since we are trying to find an \emph{injective} mapping from a pattern $P = G(p, d_p)$
to a target $T = G(t, d_t)$, there are \[ t^{\underline{p}} = t \cdot (t - 1) \cdot \ldots \cdot (t -
p + 1) \] possible assignments of target vertices to pattern vertices.  We expect the pattern to have
$d_p \cdot \binom{p}{2}$ edges, so we obtain the probability of all of these edges being mapped to
edges in the target by raising $d_t$ to this power, giving an expected number of solutions of \[
\langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \textnormal{.} \] This
formula predicts a very sharp phase transition from $\langle Sol \rangle \ll 1$ to $\langle Sol
\rangle \gg 1$, which may easily be located numerically. We plot where this occurs using black lines
in the first row of \cref{figure:non-induced}.

@@

To predict the location of the induced phase transition, we repeat the argument for locating the
non-induced phase transition and additionally considering non-edges, to get an expected number of
solutions of \[ \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot
{(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \textnormal{.} \] We plot this using black lines on
the top row of \cref{figure:induced}---again, our prediction is accurate except for very sparse or
very dense patterns.

@@

Constrainedness, denoted $\kappa$, is an alternative measure of difficulty designed to refine the
phase transition concept, and to generalise hardness parameters across different combinatorial
problems \citep{Gent:1996:Kappa}. A problem with $\kappa < 1$ is said to be underconstrained, and is
likely to be satisfiable; a problem with $\kappa > 1$ is overconstrained, and is likely to be
unsatisfiable. Empirically, problems with $\kappa$ close to 1 are hard, and problems where $\kappa$
is very small or very large are usually easy. By handling injectivity as a restriction on the size
of the state space rather than as a constraint, we derive \[ \kappa = 1 - \frac{\log \left(
t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot
\binom{p}{2}} \right)}{\log t^{\underline{p}}}\] for induced isomorphisms, which we plot on the
fifth row of \cref{figure:induced}. We see that constrainedness predicts that the central region
will still be relatively difficult for larger pattern graphs: although the problem is
overconstrained, it is less overconstrained than in the regions the Glasgow and LAD solvers found
easy.  Thus it seems that rather than just being a unification of existing generalised heuristic
techniques, constrainedness also gives a better predictor of difficulty than proximity to a phase
transition---our method generates instances where constrainedness and ``close to a phase
transition'' give very different predictions, and constrainedness gives the better prediction.

@@

As before, we expect the pattern to have $d_p \cdot \binom{p}{2}$ edges, so we obtain the
probability of all of these edges being mapped to edges in the target by raising $d_t$ to this
power, giving an expected number of solutions of \[ \langle Sol \rangle = \left(
    \frac{\Gamma\left(\nicefrac{t}{k} + 1\right)}{\Gamma\left(\nicefrac{t}{k} - \nicefrac{p}{k} +
1\right)} \right)^{k}  \cdot
{d_t}^{d_p \cdot \binom{p}{2}} \textnormal{\,.} \]

>>>>

We can approximately predict the location of the phase transition by calculating (with
simplifications regarding rounding and independence) the expected number of solutions for given
parameters. Since we are trying to find an \emph{injective} mapping from a pattern $P = G(p, d_p)$
to a target $T = G(t, d_t)$, there are \[ t^{\underline{p}} = t \cdot (t - 1) \cdot \ldots \cdot (t -
p + 1) \] possible assignments of target vertices to pattern vertices.  We expect the pattern to have
$d_p \cdot \binom{p}{2}$ edges, so we obtain the probability of all of these edges being mapped to
edges in the target by raising $d_t$ to this power, giving an expected number of solutions of
\begin{equation}
    \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \textnormal{.}
\end{equation} This
formula predicts a very sharp phase transition from $\langle Sol \rangle \ll 1$ to $\langle Sol
\rangle \gg 1$, which may easily be located numerically. We plot where this occurs using black lines
in the first row of \cref{figure:non-induced}.

@@

To predict the location of the induced phase transition, we repeat the argument for locating the
non-induced phase transition and additionally considering non-edges, to get an expected number of
solutions of \begin{equation} \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot
{(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \textnormal{.} \end{equation} We plot this using black lines on
the top row of \cref{figure:induced}---again, our prediction is accurate except for very sparse or
very dense patterns.

@@

Constrainedness, denoted $\kappa$, is an alternative measure of difficulty designed to refine the
phase transition concept, and to generalise hardness parameters across different combinatorial
problems \citep{Gent:1996:Kappa}. A problem with $\kappa < 1$ is said to be underconstrained, and is
likely to be satisfiable; a problem with $\kappa > 1$ is overconstrained, and is likely to be
unsatisfiable. Empirically, problems with $\kappa$ close to 1 are hard, and problems where $\kappa$
is very small or very large are usually easy. By handling injectivity as a restriction on the size
of the state space rather than as a constraint, we derive \begin{equation} \kappa = 1 - \frac{\log \left(
t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot
\binom{p}{2}} \right)}{\log t^{\underline{p}}} \end{equation} for induced isomorphisms, which we plot on the
fifth row of \cref{figure:induced}. We see that constrainedness predicts that the central region
will still be relatively difficult for larger pattern graphs: although the problem is
overconstrained, it is less overconstrained than in the regions the Glasgow and LAD solvers found
easy.  Thus it seems that rather than just being a unification of existing generalised heuristic
techniques, constrainedness also gives a better predictor of difficulty than proximity to a phase
transition---our method generates instances where constrainedness and ``close to a phase
transition'' give very different predictions, and constrainedness gives the better prediction.

@@

As before, we expect the pattern to have $d_p \cdot \binom{p}{2}$ edges, so we obtain the
probability of all of these edges being mapped to edges in the target by raising $d_t$ to this
power, giving an expected number of solutions of \begin{equation} \langle Sol \rangle = \left(
    \frac{\Gamma\left(\nicefrac{t}{k} + 1\right)}{\Gamma\left(\nicefrac{t}{k} - \nicefrac{p}{k} +
1\right)} \right)^{k}  \cdot
{d_t}^{d_p \cdot \binom{p}{2}} \textnormal{\,.} \end{equation}
